---
title: "Differential_expression"
author: "PaulMetz"
date: "3-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load tidyverse and DESeq

```{r}
library(tidyverse)
library(DESeq2)
```
Convert the counts table (obds_countstable.tsv.gz) and the sample
information table (obds_sampletable.tsv) into a suitable format for
generating a DESeqDataSet object

```{r}
counts_table = read.table("data/obds_countstable.tsv.gz", header = TRUE, row.names = 1)
View(counts_table)

sample_table = read.table("data/obds_sampletable.tsv", sep = "\t", header = TRUE, row.names = 1)
View(sample_table)

DESeqDataSetFromMatrix()
```

convert counts table df to matrix

```{r}
counts_table <- as.matrix(counts_table)
```

Set Egr2/3 DKO CD8 cells as the reference level
In order to do this we need to remove the replicate number from the sample title colum to be able to compare the right sample groups

```{r}
sample_table <- sample_table %>%
    separate(sample_title, into = c("egr", "ko", "cell", "replicate"))
```























































































