---
title: "sc_droplet"
author: "PaulMetz"
date: "10-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(DropletUtils)
```


Import the raw Cell Ranger matrix for 1,000 cells into R; use DropletUtils.
```{r}
sce <- read10xCounts("data/pbmc_1k_v3_raw/", sample.names = "PBMC")

View(as.data.frame(colData(sce)))
```

Remove droplets without any count at all.
```{r}
library(DropletUtils)

head(assay(sce, "counts"))
sums <- colSums(assay(sce, "counts"))
sce <- sce[, sums > 0]
sce
```



```{r}

emptyDroplets_probability.df <-  DropletUtils::emptyDrops(assay(sce, "counts"))

```

Note: How do you read the output? Which droplets would you consider empty?

```{r}
#FDR - false discovery rate. less than 0.01 pvalue considered to be a cell. 
is.cell <- emptyDroplets_probability.df$FDR <= 0.01
sum(is.cell, na.rm=TRUE)
```

Draw plots comparing empty droplets and other droplets (e.g., library size, mitochondrial content,
PCA).

```{r}
install.packages("ggplot2")
library(ggplot2)
#View(as.data.frame(is.cell))

df <- data.frame(librarysize = colSums(assay(sce, "counts")), emptydroplet = is.cell)
#View(df)

plot1 <- ggplot(df, aes(x = emptydroplet, y = librarysize)) +
  geom_point()

plot1
```





Import the 􀃘ltered Cell Ranger matrix for 1,000 cells and compare their set of 􀃘ltered cells.
```{r}

sce_filtered <- read10xCounts("data/pbmc_1k_v3_filtered/", sample.names = "PBMC")

sce_filtered

```


```{r}
head(which(is.cell))

sce <- sce[,which(is.cell)]
sce
```
compare how many cells are overlapping between sce and sce_filtered

```{r}
barcodes_raw <- colData(sce)$Barcode
barcodes_filtered <- colData(sce_filtered)$Barcode

length(intersect(barcodes_raw, barcodes_filtered))
```
so above we see that 1773 cells overlap, so we did a similar filtering out of empty droplets. 


#removing doublets
```{r}
BiocManager::install("scDblFinder")
library(scDblFinder)

sce <- scDblFinder(sce)


View(as.data.frame(colData(sce)))


```


```{r}
df2 <- data.frame(librarysize = colSums(assay(sce, "counts")), doublets = colData(sce)$scDblFinder.class)
View(df2)

plot2 <- ggplot(df2, aes(x = doublets, y = librarysize)) +
  geom_violin()+
  geom_point()

plot2
```

```{r}
head(colData(sce)$scDblFinder.class == "singlet")

sce <- sce[,colData(sce)$scDblFinder.class == "singlet"]
sce
```

```{r}
BiocManager::install("celda")

```


```{r}
library(celda)

sce <- decontX(sce)
assays(sce)$decontXcounts
```



Visualize the UMAP layout created by decontX ; use plotDimReduceCluster() .
```{r}
View(as.data.frame(colData(sce)))

reducedDim(sce)

class(sce)

plotDimReduceCluster(colData(sce)$decontX_clusters, dim1 = reducedDim(sce)[,"DecontX_UMAP_1"], dim2= reducedDim(sce)[,"DecontX_UMAP_2"])
```


Visualize the percentage of contamination estimated by decontX in each cell. Use the
plotDecontXContamination() function.
```{r}
plotDecontXContamination(sce)
```
Visualize the log-normalised expression of marker genes on a UMAP layout faceted by gene; use
plotDimReduceFeature().

```{r}

reducedDimNames(sce)

head(rownames(sce))

plotDimReduceFeature(as.matrix(assays(sce)$counts), features = "ENSG00000198851", dim1 = reducedDim(sce)[,"DecontX_UMAP_1"], dim2= reducedDim(sce)[,"DecontX_UMAP_2"])

```

Visualise the percentage of cells with detectable expression of marker genes before and after
decontamination as a bar plot; use plotDecontXMarkerPercentage() .

```{r}
plotDecontXMarkerPercentage(sce, markers)

```




































