---
title: "scrnaseq"
author: "PaulMetz"
date: "8-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries

```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
```

load in data and subsetting

```{r}
pbmc.data <- Read10X(data.dir = "filtered_feature_bc_matrix/")
pbmc <- CreateSeuratObject(counts = pbmc.data[["Gene Expression"]], project = "pbmc3k")
```

using this website: https://satijalab.org/seurat/articles/pbmc3k_tutorial.html or this one https://satijalab.org/seurat/articles/multimodal_vignette.html


add the protein data matrix as an assay to the pbmc seurat matrix we just created
```{r}

# create a new assay to store ADT information
adt_assay <- CreateAssayObject(counts = pbmc.data[["Antibody Capture"]])

# add this assay to the previously created Seurat object
pbmc[["ADT"]] <- adt_assay
```

accessing slots from pbmc, e.g.:
```{r}
#example
data_RNA <- FetchData(object = pbmc, vars = "ident", slot = "data")
#head(data_RNA)
#View(pbmc[[]]) #this SOMEHOW views the metadata slot 
```


```{r}
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-") #looks for all genes starting with ^MT- and calculates percentage of counts of mt dna genes in all counts and puts it into a new colum
#View(pbmc[[]])
```
# Visualize QC metrics as a violin plot
```{r}

VlnPlot(pbmc, features = "nFeature_RNA", combine = TRUE) + geom_hline(yintercept = c(1000, 5000))
VlnPlot(pbmc, features = "percent.mt", combine = TRUE) + geom_hline(yintercept = c(15))


#VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
#QC plot of nCount_RNA by nFeature_RNA coloured by percent.mt (use ggplot2)

```{r}
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot3 <- ggplot(pbmc[[]], aes(x = nCount_RNA, y = nFeature_RNA, col = percent.mt)) +
    geom_point()

#plot1 + plot2 + plot3
```

based on above plots, we will set the following treshholds for filtering:

# pbmc <- subset(pbmc, subset = nFeature_RNA > 1000 & nFeature_RNA < 5000 & percent.mt < ?)


# run sctransform

```{r}
pbmc <- SCTransform(pbmc, vars.to.regress = "percent.mt", verbose = FALSE)
#you will get a lot of warnings. 
#View(pbmc)
```



```{r}
DefaultAssay(pbmc)
DefaultAssay(pbmc) <- "SCT" #Change default
DefaultAssay(pbmc)
```

```{r}
pbmc <- RunPCA(pbmc, npcs = 50) #npcs = 50 by default, we can change this. 
```

```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc, reduction = "pca", dims = c(1,2))
```


```{r}
ElbowPlot(pbmc, ndims = 50)
```

Clustering

```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:20, k.param = 20)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```



```{r}
pbmc <- RunUMAP(pbmc, dims = 1:20)
DimPlot(pbmc, reduction = "umap", label = TRUE)
```


```{r}
FeaturePlot(pbmc, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP", 
    "CD8A"))
```

Generate new plot that shows mitochondrial expression, to see which have high vs low expression


```{r}
FeaturePlot(pbmc, features = "percent.mt")
FeaturePlot(pbmc, features = "nFeature_RNA", min.cutoff = 1000, max.cutoff = 5000)


DimPlot(pbmc, cells.highlight = WhichCells(pbmc, expression = percent.mt < 15)) + NoLegend()

```

```{r}
#here we get rid of clusters we dont want to keep:
pbmc_subset <- subset(pbmc, idents = c("1", "10"), invert = TRUE)

DimPlot(pbmc_subset)


```


```{r}
DimPlot(pbmc_subset, cells.highlight = WhichCells(pbmc_subset, expression = nFeature_RNA > 5000)) + NoLegend()

DimPlot(pbmc_subset, cells.highlight = WhichCells(pbmc_subset, expression = nFeature_RNA < 1000)) + NoLegend()
```



☺
```{r}
pbmc_subset <- SCTransform(pbmc_subset, vars.to.regress = "percent.mt", verbose = FALSE)
pbmc_subset <- RunPCA(pbmc_subset, pncs = 50)
pbmc_subset <- FindNeighbors(pbmc_subset, dims = 1:20, k.param = 20)
pbmc_subset <- FindClusters(pbmc_subset, resolution = 0.5)
pbmc_subset <- RunUMAP(pbmc_subset, dims = 1:20)
DimPlot(pbmc_subset, reduction = "umap", label = TRUE) 

FeaturePlot(pbmc_subset, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP", 
    "CD8A"))
```


```{r}
DimPlot(pbmc_subset, reduction = "umap", label = TRUE)
```




















































