---
title: "Classification"
author: "PaulMetz"
date: "2-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load caret

```{r}
library(caret)
```

Get familiar with the data set. What are rows? What are columns?

```{r}
View(iris)
class(iris)
```

Use the caret::featurePlot() function to visualise species and measurements. Try the different options for the plot argument.

```{r}
featurePlot(x = iris[,1:4], y = iris$Species, plot = "box" )
```

Partition the data set into one training and one test set. Make sure the two sets are balanced with respect to Species .

```{r}
set.seed(998)
inTraining <- createDataPartition(iris$Species, p = .75, list = FALSE)
training <- iris[ inTraining,]
testing <- iris[-inTraining,]
```

```{r}
table(training$Species)
```
Train a model, e.g. linear SVM (Support Vector Machine), random forest.

```{r}
install.packages("party")
install.packages("e1071")

fitControl <- trainControl(
    method = "boot", 
    number = 5) ## repeated five times

modelLookup("cforest")

mtryFit <- train(
    Species ~ ., data = training,
    method = "cforest", trControl = fitControl,
    tuneGrid = data.frame(mtry = c(1,2,3,4)))

mtryFit
```

graph

```{r}
ggplot(mtryFit)
```




Use the model to make predictions on the test set. How accurate is the model?

```{r}
cforestPred <- predict(mtryFit, newdata = testing)
confusionMatrix(testing$Species, data = cforestPred)$table
class(testing$Species)
str(testing$Species)
class(cforestPred)
str(testing$Species)
```

```{r}
confusionMatrix(testing$Species, data = cforestPred)$overall["Accuracy"]
```
















