---
title: "linear_regression_machine_learning"
author: "PaulMetz"
date: "2-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load tidyverse

```{r}
library(tidyverse)
```

The ChickWeight data set measures the impact of different diets on the early growth of chicks.
Fit a linear mode to measure the effect of Time and Diet in the ChickWeight data set.

```{r}
View(ChickWeight)
chicks_lm_fit <- lm(weight ~ Time + Diet, ChickWeight)

chicks_lm_fit
```
summary(chicks_lm_fit)

```{r}
summary(chicks_lm_fit)
```

plot the linear model

```{r}
plot1 <- ggplot(ChickWeight, aes(Time, weight, col = Diet)) +
    geom_point() +
    stat_smooth(method = "lm")

plot1
```

Linear model for individual variables

```{r}
chicks_lm_fit <- lm(weight ~ Time, ChickWeight)

chicks_lm_fit
```
plot the new linear model

```{r}
plot2 <- ggplot(ChickWeight, aes(Time, weight)) +
    geom_point() +
    stat_smooth(method = "lm") +
    geom_abline(slope=8.803, intercept = 27.467)

plot2
```


```{r}
chicks_lm_fit <- lm(weight ~ Diet, ChickWeight)

chicks_lm_fit
```
```{r}
plot3 <- ggplot(ChickWeight, aes(Diet, weight)) +
    geom_boxplot() +
    geom_point() 

plot3
```



So from above we can see that each diet has a different slope(first plot). So the linear model we made before doesn't correlate with the lm graph we made. So lets improve the lm model. 

```{r}
chicks_lm_fit <- lm(weight ~ Time * Diet, ChickWeight)

chicks_lm_fit
```
See above: intercept: 30.9; where diet1 intercepts x = 0. Time:Diet2 says how much influence time has on the weight for diet2, compared to diet1. They all take Diet1 as a base so to say. 




plot the linear model again

```{r}
plot4 <- ggplot(ChickWeight, aes(Time, weight, col = Diet)) +
    geom_point() +
    stat_smooth(method = "lm") +
    geom_abline(slope = 6.8418 + 4.5811, intercept = 30.9310 - 12.6807)
plot4
```

How much does weight increase per unit of Time for the top diet?

so that would be 6.8418 + 4.5811 = 11.4229

Does the top diet drive an increase in body weight that is significantly faster than the next best diet?

```{r}
chicks_lm_fit <- lm(weight ~ Time * Diet, ChickWeight)
summary(chicks_lm_fit)

```

above doesnt show the comparison between diet3 and diet4, which are the top two diets. so we need to compared those instead of diet3 to diet1.

```{r}
ChickWeight$Diet <- relevel(ChickWeight$Diet, "4")
chicks_lm_fit <- lm(weight ~ Time * Diet, ChickWeight)
summary(chicks_lm_fit)
```

predict based on value
```{r}
predict(chicks_lm_fit, tibble(Time = 5, Diet = "3"))
```












