---
title: "Regression_excersise"
author: "PaulMetz"
date: "2-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load caret

```{r}
library(caret)
install.packages("e1071")
library(Cubist)
```

load Boston housing data

```{r}
library(mlbench)
data(BostonHousing)
```

Get familiar with the data set. What are rows? What are columns?

```{r}
View(BostonHousing)
```

Partition the data set into one training and one test set. Make sure the two sets are balanced with respect to medv (the median house value of districts).

```{r}
set.seed(998)
inTrain <- sample(1:nrow(BostonHousing), floor(.8*nrow(BostonHousing)))

train_pred <- BostonHousing[ inTrain, -14]
test_pred  <- BostonHousing[-inTrain, -14]

train_resp <- BostonHousing$medv[ inTrain]
test_resp  <- BostonHousing$medv[-inTrain]

View(train_resp)
```

Train a model, e.g. a linear regression model.
Inspect and evaluate the performance of the model during training.

```{r}
#modelLookup("cubist")
model_tree <- cubist(x = train_pred, y = train_resp, committees = 100)
summary(model_tree)
```
#https://cran.r-project.org/web/packages/Cubist/vignettes/cubist.html 
Use the model to make predictions on the test set. How accurate is our model?

```{r}

model_tree_pred <- predict(model_tree, test_pred, neighbors = 2)
cor(model_tree_pred, test_resp)^2

```




























