---
title: "bioconductor_scsequencing"
author: "PaulMetz"
date: "9-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

http://bioconductor.org/books/release/OSCA/overview.html#quick-start


alternative to seurat, both are good. depends on question. 

load library
```{r}
library(DropletUtils)
```

1. Import the 􀃘ltered matrix into R; use DropletUtils. What can you tell about the raw data?

```{r}
sce <- read10xCounts("data/", sample.names = )
#View(sce)
#dim(sce)
#str(sce)

temp <- as.data.frame(colData(sce))
length(temp)

#create a sequence of number sfrom 1 to number of columns in sce
sce <- read10xCounts("data/", sample.names = "pbmc")

#View(as.data.frame(colData(sce)))
#dim(sce)
```

2. Compute and visualise quality control metrics; use scatter. Would you remove any cell?

```{r}
library(scater)
perCellQCMetrics(sce)
    #subsets = NULL, later we can change this to subset/filter mitochondrial genes expressing cells. 

sce <- addPerCellQC(sce)
#View(as.data.frame(colData(sce)))

rowData(sce) #ensemble IDs
rowData(sce) #symbol colum showing gene names

# Mitochondrial genes
mito_genes <- rownames(sce)[grep("^MT-", rowData(sce)$Symbol)]

# Ribosomal genes
ribo_genes <- rownames(sce)[grep("^RP[SL]", rowData(sce)$Symbol)]

# Hemoglobin genes - includes all genes starting with HB except HBP.
hb_genes <- rownames(sce)[grep("^HB[^(P)]", rowData(sce)$Symbol)]

sce2 <- addPerCellQC(sce, subsets = list(mt = mito_genes, hb = hb_genes, 
    ribo = ribo_genes))

#View(as.data.frame(colData(sce2)))
```

#visialise quality metrics

```{r}
plot1 <- ggplot(as.data.frame(colData(sce2)), aes(x = subsets_mt_percent))+geom_histogram(bins= 500)+geom_vline(xintercept = 20)
```
#see above, probably cut of at 20% or so. 

```{r}
plot2 <- ggplot(as.data.frame(colData(sce2)), aes(x = sum))+geom_histogram(bins = 100)+geom_vline(xintercept = c(3000, 20000))
```

```{r}
plot3 <- ggplot(as.data.frame(colData(sce2)), aes(x = detected))+geom_histogram(bins = 100)+geom_vline(xintercept = c(500, 2500))
```

```{r}
plot4 <- ggplot(as.data.frame(colData(sce2)), aes(x = sum, y = detected))+geom_point()
```

```{r}
library(cowplot)
plot_grid(plot1, plot2, plot3, plot4)
```

```{r}
sce <- sce2
dim(sce)

sce_subset <- sce[,colData(sce)$subsets_mt_percent < 15 & colData(sce)$sum > 2500 & colData(sce)$sum < 20000 & colData(sce)$detected > 1000 & colData(sce)$detected < 4250]

dim(sce_subset)
```
```{r}
perFeatureQCMetrics(sce_subset)

sce_subset <- addPerFeatureQC(sce_subset)


sce_subset <- sce_subset[rowData(sce_subset)$detected >0,]
#dim(sce_subset)
```

Convert the counts into normalized expression values to eliminate cell-speci􀃘c biases (e.g., in capture
ef􀃘ciency); use scater and/or scran.

```{r}
library(scran)
#View(sce)
sce_subset <- computeSumFactors(sce_subset)
sce_subset <- logNormCounts(sce_subset)
assayNames(sce_subset)
```
Plot the variance against the mean of each gene.
```{r}
#View(sce_subset)

x <- DelayedArray(assay(sce_subset, "counts"))
mean <- DelayedMatrixStats::rowMeans2(x)
variance <- DelayedMatrixStats::rowVars(x)
df <- data.frame(mean = mean, variance = variance)

plot5 <- ggplot(df, aes(x = log2(mean+1), y = log2(variance+1))) +
    geom_point()
plot5

```

```{r}
#View(sce_subset)

x <- DelayedArray(assay(sce_subset, "logcounts"))
mean <- DelayedMatrixStats::rowMeans2(x)
variance <- DelayedMatrixStats::rowVars(x)
df <- data.frame(mean = mean, variance = variance)

plot6 <- ggplot(df, aes(x = mean, y = variance)) +
    geom_point()
plot6

```

Select features for downstream analyses, e.g. highly variable genes; use scran.
Use scran::modelGeneVar() to model the variance of the log-expression pro􀃘les for each gene.
What is the output?
Visualise the relation between the mean expression of each gene and the total / biological / technical variance of each gene.


```{r}
gene_variance.df <- scran::modelGeneVar(sce_subset)
#View(gene_variance.df)

gene_variance.df <- as.data.frame(gene_variance.df)

#View(as.data.frame(gene_variance.df))

plot7 <- ggplot(gene_variance.df, aes(x = mean, y = total)) +
    geom_point()
plot8 <- ggplot(gene_variance.df, aes(x = mean, y = tech)) +
    geom_point()
plot9 <- ggplot(gene_variance.df, aes(x = mean, y = bio)) +
    geom_point()

plot_grid(plot7, plot8, plot9)

```
Note: How do you interpret those different values?
Use scran::getTopHVGs() to identify highly variable genes.
Note: what is the output? How many genes do you identify? Where are those genes located in the meanvariance
plot? What happens to this plot if you set more stringent thresholds to de􀃘ne highly variable
genes?

```{r}
top10_most_variable_genes <- scran::getTopHVGs(gene_variance.df, var.field = "bio", prop = 0.1)
```

```{r}
set.seed(5637)
sce_subset <- scater::runPCA(sce_subset, subset_row = top10_most_variable_genes, ncomponents = 50)

percent.var <- attr(reducedDim(sce_subset), "percentVar")

#View(percent.var)

df2 <- data.frame(x = seq(1, 50), y = percent.var)

plot10 <- ggplot(df2, aes(x = x, y = y)) +
    geom_col()

plot10
#reduced(sce_subset)
#reducedDimNames(sce_subset)

```
umap
updated to use 12 dimensions since we found this later on. 
```{r}
sce_subset <- scater::runUMAP(sce_subset, pca = 12)

reducedDimNames(sce_subset)
```
Visualise the layout of cells produced by each of those dimensionality reduction methods.
Considering coloring points with quality control metrics

```{r}
#View(sce_subset)

plotReducedDim(sce_subset, dimred = "UMAP", colour_by = "subsets_mt_percent")

```

```{r}
#View(sce_subset)

plotReducedDim(sce_subset, dimred = "PCA", colour_by = "subsets_mt_percent")

```

```{r}
plotPCA(sce_subset, ncomponents = 4, colour_by = "subsets_mt_percent")
```

Exercise step 6. Clustering

Use scran::getClusteredPCs() to cluster cells after using varying number of PCs and pick the
number of PCs using a heuristic based on the number of clusters.

```{r}

#View(sce_subset)

PCs_matrix <- reducedDim(sce_subset)

clustered_PCs.df <- scran::getClusteredPCs(PCs_matrix)


metadata(clustered_PCs.df)$chosen #tells us 12, so that should be the ideal number of clusters. 

```

Use scran::buildSNNGraph() and igraph::cluster_louvain() with the "ideal" number of PCs.

d = 12 based on previous
maybe change k depending on what you get.

```{r}
library(igraph)
snngraph <- scran::buildSNNGraph(sce_subset, d = 12)

```

we need to look at membership which tells us which cluster is used for each cell.

```{r}
louvain <- igraph::cluster_louvain(snngraph)
length(membership(louvain))
cluster_identity <- as.factor(membership(louvain)) # need to change to numeric to be able to use later on
dim(sce_subset)
```
Visualise the assigned cluster on your preferred dimensionality reduction layout.

we made UMAP already, lets make TSNE to compare

```{r}
sce_subset <- scater::runTSNE(sce_subset, pca = 12)

reducedDimNames(sce_subset)
```
```{r}
#we need to add cluster_identity to colData in the sce_subset object.

colData(sce_subset)$cluster_identity <- cluster_identity
```

now we can plot
```{r}
View(colData(sce_subset))

plotReducedDim(sce_subset, dimred = "UMAP", colour_by = "cluster_identity", text_by = "cluster_identity")

```

now plot the TSNE one


```{r}
plotReducedDim(sce_subset, dimred = "TSNE", colour_by = "cluster_identity", text_by = "cluster_identity")
```

Exercise step 7. Cluster markers

```{r}


markers <- scran::findMarkers(sce_subset, groups = cluster_identity, test.type = "wilcox")

#markers[[1]]
```

Visualise the expression of selected markers:
as a dot plot, optionally with a violin layer.
on a dimensionality reduction layout.


```{r}
rownames(sce_subset) <- rowData(sce_subset)$Symbol
scater::plotUMAP(sce_subset, colour_by = "S100A4")
```
```{r}
plotExpression(sce_subset, features=c("S100A4", "TRAC"),
x="cluster_identity", colour_by="cluster_identity")
```



http://bioconductor.org/books/release/OSCA/overview.html#quick-start








































