---
title: "Dimensionality_reduction"
author: "PaulMetz"
date: "1-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in data

```{r}
logcounts <- read.csv("data/logcounts.csv", row.names = 1)
View(logcounts)

```

Load tidyverse

```{r}
library(tidyverse)
```

Convert to a matrix and transpose

```{r}
logcounts_matrix_transposed <- t(as.matrix(logcounts))
View(logcounts_matrix_transposed)
```

Perform PCA. Inspect the structure of the output.

```{r}

logcounts.pca <- prcomp(logcounts_matrix_transposed, center = TRUE, scale = FALSE)
str(logcounts.pca)
logcounts_pca_scores <- as.data.frame(logcounts.pca$x)
```

Draw a scatter plot of the top two principal components and color by various experimental metadata in cell_metadata.csv. Which experimental factors show the largest variance in the data?

```{r}
cell_metadata <- read.csv("data/cell_metadata.csv")
View(cell_metadata)
```

Select first two columns for above exercise

```{r}
first_two_pcas <- logcounts_pca_scores[,1:2]
View(first_two_pcas)
```

Convert rownames to a column

```{r}
table_for_plotting <- first_two_pcas %>% 
    rownames_to_column(var = "Sample") %>%
    full_join(cell_metadata)

View(temp)    
        
```

Make scatterplot

```{r}

plot1 <- ggplot(table_for_plotting, aes(x = PC1, y = PC2, col = Status)) +
    geom_point()

plot1
```

Draw a scatter plot of the top two principal components and color by various experimental metadata
in cell_metadata.csv . Which experimental factors show the largest variance in the data?

```{r}

plot2 <- ggplot(table_for_plotting, aes(x = PC1, y = PC2, col = Infection)) +
    geom_point()

plot3 <- ggplot(table_for_plotting, aes(x = PC1, y = PC2, col = Status)) +
    geom_point()

plot4 <- ggplot(table_for_plotting, aes(x = PC1, y = PC2, col = Time)) +
    geom_point()

library(cowplot)

plot_grid(plot2, plot3, plot4, labels = c("Infection", "Status", "Time"), ncol = 1, nrow = 3)


```

Draw a density plot, faceted by time and infection, colored by status, to inspect the variance of each
experimental factor on the top principal component.

```{r}
plot5 <- ggplot(table_for_plotting, aes(x = PC1, col = Status)) +
    geom_density() +
    facet_grid(rows = vars(Infection), cols = vars(Time))

plot5
```

Draw a bar plot of variance explained by each principal component. How many principal components
would you keep for downstream analyses?


```{r}
variance_table <- data.frame(variance = (logcounts.pca$sdev)^2)
variance_table$fraction <- variance_table$variance/sum(variance_table$variance)
PC_numbers <- seq(1,nrow(variance_table))
variance_table$PC <- PC_numbers
variance_table <- variance_table %>%
    select(PC, variance, fraction)
View(variance_table)

plot6 <- ggplot(variance_table, aes(x = PC, y = fraction)) +
    geom_col() +
    xlim(0, 100)

plot6

```

Find the names of the top genes associated with the top principal components.

cumulative sum

```{r}

variance_table$cum_sum <- cumsum(variance_table$fraction)

View(variance_table)

plot7 <- ggplot(variance_table, aes(x = PC, y = cum_sum)) +
    geom_line() +
    xlim(0, 100)

plot7

```

Find the names of the top genes associated with the top principal components.


```{r}
genes_influence <- as.data.frame(logcounts.pca$rotation)
genes_influence$gene_ID <- rownames(genes_influence)

temp <- genes_influence %>%
    arrange(desc(PC1)) %>%
    select(gene_ID, PC1)

View(temp)
```


Visualise the gene expression value of the gene with the highest loading for PC1 (ENSG00000172183) as color on a scatter plot of PC1 and PC2. What do you observe?


```{r}
View(table_for_plotting)
ENSG00000172183 <- as.numeric(logcounts["ENSG00000172183",])
ENSG00000172183

table_for_plotting$ENSG00000172183 <- ENSG00000172183
View(table_for_plotting)

plot8 <- ggplot(table_for_plotting, aes(x = PC1, y = PC2, col = ENSG00000172183)) +
    geom_point()

plot8
```

Compare PCA to t-SNE to UMAP. Note that t-SNE and UMAP should be given the subset of principal
components selected above.


```{r}
library(umap)

subset_logcounts.pca <- logcounts.pca$x[,1:15]

View(subset_logcounts.pca)
View(cell_metadata)


logcounts_umap <- umap(subset_logcounts.pca)
logcounts_umap_layout <- as.data.frame(logcounts_umap$layout)
logcounts_umap_layout$Sample <- rownames(logcounts_umap_layout)
View(logcounts_umap_layout)

logcounts_umap_layout_metadata <- logcounts_umap_layout %>%
    full_join(cell_metadata)
View(logcounts_umap_layout_metadata)

plot9 <- ggplot(logcounts_umap_layout_metadata, aes(x = V1, y = V2, col = Time)) +
    geom_point()

plot9

```


```{r}
out_kmeans <- kmeans(x = logcounts.pca$x[, 1:25], centers = 4)
str(out_kmeans)
```



```{r}
kmeans_scree <- lapply(seq_len(30), function(x) {
    out <- kmeans(x = logcounts.pca$x[, 1:25], centers = x)
    tibble(
        k = x,
        totss = out$totss,
        betweenss = out$betweenss,
        withinss_sum = sum(out$withinss)
    )
}) %>%
    bind_rows()

kmeans_scree


```



```{r}
ggplot(kmeans_scree, aes(k, withinss_sum)) +
geom_line(linetype = "F1", size = 0.2) +
geom_point() +
labs(title = "sum(withinss)") +
cowplot::theme_cowplot()


```


```{r}
comparison_table <- tibble(
sample = rownames(logcounts.pca$x),
cell_metadata[rownames(logcounts.pca$x), ] %>% as_tibble(),
kmeans = kmeans(x = logcounts.pca$x, centers = 4)$cluster,
)
table(comparison_table$Time, comparison_table$kmeans)
```












