---
title: "integration_scseq"
author: "PaulMetz"
date: "11-3-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Import the two RDS files into R

```{r}
install.packages("remotes")
install.packages("rlang")
#packageVersion("rlang")
install.packages("Seurat")
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)

pbmcv2_filtered <- readRDS("data/pbmcv2_filtered.rds")
pbmcv3_filtered <- readRDS("data/pbmcv3_filtered.rds")
```

2. Create a merged Seurat object containing both sets of data

```{r}
pbmc.combined <- merge(x = pbmcv2_filtered, y = pbmcv3_filtered, add.cell.ids = c("v2", "v3"), project = "pbmc" )
```


3. On the merged Seurat object, perform the general steps of a Seurat analysis workflow:
Normalisation
Finding variable genes
Scaling
PCA
Clustering
UMAP

```{r}
pbmc.combined <- SCTransform(pbmc.combined, vars.to.regress = "percent.mt", verbose = FALSE)

VlnPlot(pbmc.combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.1)
```



```{r}
FeatureScatter(pbmc.combined, feature1 = "nCount_RNA", feature2 = "percent.mt")
```



4. Visualise the identified clusters on the UMAP

```{r}
pbmc.combined <- RunPCA(pbmc.combined)

```
```{r}
ElbowPlot(pbmc.combined, ndims = 50)
```
Maybe based on above we can take 20 PCs for umap.

5. Colour the cells in the UMAP by the original dataset from which they came i.e. v2 or v3
Do you see any batch effects?

now we 
```{r}
#findneighbours is greating graphs of nearest neigbours
pbmc.combined <- FindNeighbors(pbmc.combined, dims = 1:20, k.param = 20)

#split the graph into groups of cells that are similar to eachother.
pbmc.combined <- FindClusters(pbmc.combined, resolution = 0.5) #by default uses Louvain
```

```{r}
pbmc.combined <- RunUMAP(pbmc.combined, dims = 1:20)

DimPlot(pbmc.combined, reduction = "umap", label = TRUE, group.by = "orig.ident")
```

```{r}
FeaturePlot(pbmc.combined, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP", "CD8A"))
```

```{r}
pbmc.combined <- FindClusters(pbmc.combined, resolution = seq(0.1, 1, 0.1)) #you can see this in metadata
```


```{r}
library(clustree)
clustree(pbmc.combined)
#at 0.0 it was divided into three clusters, at .1 into 6 clusters, 0.2 into 8 clusters. etc. 
```


so we have seen a clear batch effect, so we'll use suerat integration to improve this ( we use the vignette for version 3). 

https://satijalab.org/seurat/archive/v3.2/integration.html 


now we need to transform our separate datasets before integration.
```{r}
pbmcv2.t <- SCTransform(pbmcv2_filtered, vars.to.regress = "percent.mt", verbose = FALSE)

pbmcv3.t <- SCTransform(pbmcv3_filtered, vars.to.regress = "percent.mt", verbose = FALSE)
```




```{r}
DefaultAssay(pbmcv2.t)
DefaultAssay(pbmcv3.t)
```

we then want to select features that are common between both datasets
```{r}
# we need list as input for next function
pbmc.list <- list(pbmcv2.t, pbmcv3.t)

#here we use the list
pbmc.features <- SelectIntegrationFeatures(object.list = pbmc.list, nfeatures = 3000)

#View(pbmc.features) #we see a vector of gene names (features)

#we need two variables that contain the variable features from each data set, not necessary
pbmcv2_features <- VariableFeatures(pbmcv2.t)
pbmcv3_features <- VariableFeatures(pbmcv3.t)
#intersect(pbmcv2_features, pbmcv3_features)
length(intersect(pbmcv2_features, pbmcv3_features))

pbmc.list <- PrepSCTIntegration(object.list = pbmc.list, anchor.features = pbmc.features, 
    verbose = FALSE)

#View(pbmc.list)
```
we can use a venn daigram to visualise the overlap
```{r}
install.packages("VennDiagram")
library(ggplot2)
library(VennDiagram)

venn <- venn.diagram(x = list(pbmcv2_features, pbmcv3_features),
category.names = c("pbmcv2", "pbmcv3"), filename = NULL,
output = TRUE)

grid.draw(venn)
```


1. Integrate the two datasets using Seurat v3 (following the vignette)
Plot a UMAP of the integrated data coloured by batch
Perform clustering on the integrated dataset

```{r}
#prepares both objects to have the same features, so we can later integrate both objects. 
pbmc.list <- PrepSCTIntegration(object.list = pbmc.list, anchor.features = pbmc.features)

#check if rownames is pbmc objects in list are the same
#before
rownames_pbmcv4 <- rownames(GetAssayData(object = pbmcv2.t, slot = "scale.data"))
rownames_pbmcv5 <- rownames(GetAssayData(object = pbmcv3.t, slot = "scale.data"))
length(intersect(rownames_pbmcv4, rownames_pbmcv5))
identical(rownames_pbmcv4, rownames_pbmcv5)

#after
rownames_pbmcv2 <- rownames(GetAssayData(object = pbmc.list[[1]], slot = "scale.data"))
rownames_pbmcv3 <- rownames(GetAssayData(object = pbmc.list[[2]], slot = "scale.data"))

length(intersect(rownames_pbmcv2, rownames_pbmcv3))
identical(rownames_pbmcv2, rownames_pbmcv3)
```

```{r}

pbmc.anchors <- FindIntegrationAnchors(object.list = pbmc.list, normalization.method = "SCT", 
    anchor.features = pbmc.features, verbose = FALSE, dims = 1:30)  #integration can be optimised by changing dims. 

overlapping_features_pbmcobjectsv2 <- rownames(pbmc.list[[1]])
overlapping_features_pbmcobjectsv3 <- rownames(pbmc.list[[2]])


pbmc.integrated <- IntegrateData(anchorset = pbmc.anchors, normalization.method = "SCT", 
    verbose = FALSE, dims = 1:30, features.to.integrate = intersect(overlapping_features_pbmcobjectsv2, overlapping_features_pbmcobjectsv3)) #use same number of dims

DefaultAssay(pbmc.integrated) #default assay to work on is integrated

```

run PCA and UMAP
```{r}
pbmc.integrated <- RunPCA(pbmc.integrated, verbose = FALSE)

#ElbowPlot(pbmc.integrated, ndims = 50) #to decide how many PCs we will use for UMAP

pbmc.integrated <- RunUMAP(pbmc.integrated, dims = 1:20)
```

```{r}
#View(pbmc.integrated[[]]) #check which column we need to group by
plot1 <- DimPlot(pbmc.integrated, group.by = "orig.ident")
plot1
```
```{r}
pbmc.integrated <- FindNeighbors(pbmc.integrated, dims = 1:20, k.param = 20)
pbmc.integrated <- FindClusters(pbmc.integrated, resolution = 0.5)

#Reductions(pbmc.integrated) # to see which dimension reductions you have

DimPlot(pbmc.integrated, reduction = "umap", label = TRUE) 
```

```{r}
FeaturePlot(pbmc.integrated, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP", "CD8A"))
```

```{r}
install.packages("devtools") 
devtools::install_github("TomKellyGenetics/leiden")

library(leiden)

install.packages("reticulate")
reticulate::py_install(packages ='leidenalg')
reticulate::py_install(packages ='igraph')
reticulate::py_install(packages ='pandas')

#lets use a different algorithm to see if it's giving different clustering
leiden_pbmc.integrated <- FindClusters(pbmc.integrated, resolution = 0.5, algorithm = 4)

#View(leiden_pbmc.integrated[[]]) #[[]] to view metadata

leiden_pbmc_integrated <- FindClusters(leiden_pbmc.integrated, resolution = 0.8, algorithm = 4) # clustering with Leiden

colnames(leiden_pbmc_integrated@meta.data)[9] <- "Leiden_0.8" #changing the name of the metadata column with the Leiden clustering

leiden_pbmc_integrated <- FindClusters(leiden_pbmc_integrated, resolution = 0.8) # clustering with Louvain at the same resolution
#View(pbmc_integrated[[]])

DimPlot(leiden_pbmc_integrated, reduction = "umap", label=TRUE, group.by="Leiden_0.8")

DimPlot(leiden_pbmc_integrated, reduction = "umap", label=TRUE, group.by="integrated_snn_res.0.8")
```

for harmony: 

```{r}

#below code may not work, may need additional packages and changing variable names, as you copied you it from someone else. 
BiocManager::install("MAST")
library(MAST)

#change default assay from integrated to RNA
DefaultAssay(pbmc.integrated) <- "RNA"

DefaultAssay(pbmc.integrated)

DE_genes <- FindAllMarkers(pbmc.integrated, test.use = "MAST", latent.vars = "orig.ident")
Idents(pbmc_integrated) <- "SCT_snn_res.0.5"
```































