---
title: "Data analysis using Tidyverse"
author: "Oxford Biomedical Data Science Training Programme"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Prepare data

### Read data in - `read_csv()`

```{r}
raw_data <- read_csv("data/140915_counts_rfp_pos_single_cell2.csv")
```

### `pivot_longer()` - convert to long format

```{r}
test_with_na <- raw_data %>% 
    pivot_longer(-crypt_no, names_to = "mouse_typeofcount", values_to = "count")         # This will have NAs for crypt numbers that weren't counted for a given mouse

data_processed <- raw_data %>% 
    pivot_longer(-crypt_no, names_to = "mouse_typeofcount", values_to = "count", 
                 values_drop_na = TRUE)
```

### `separate()` mouse and count type

```{r}
data_processed <- data_processed %>% 
    separate(mouse_typeofcount, c("Mouse_ID", "count_type"), sep = "_")
```

### `pivot_wider()` - want two columns (RFPpos, TotalCells)

```{r}
data_processed <- data_processed %>% 
    pivot_wider(names_from = count_type, values_from = count)
```

### Do all the processing steps together

```{r}
data_processed <- raw_data %>% 
    pivot_longer(-crypt_no, names_to = "mouse_typeofcount", values_to = "count", 
                 values_drop_na = TRUE) %>% 
    separate(mouse_typeofcount, c("Mouse_ID", "count_type"), sep = "_") %>% 
    pivot_wider(names_from = count_type, values_from = count)     
```

### Join mouse info

```{r}
mouse_info <- read_csv("data/mouse_info.csv")
data_processed <- data_processed %>% 
    left_join(mouse_info)
# Will join by one or more column names if the same in both tables
# Sometimes you need to specify the columns to match e.g. by = c("from_name" = "to_name")

# Data is now tidy!!
```

### Rename and reorder columns

```{r}
data_processed <- data_processed %>%
    rename(Crypt_number = crypt_no, Total_cells = TotalCells, 
           Day_post_label = Day_PostLabel) %>%
    select(Mouse_ID, Sex, Crypt_number, Day_post_label, RFPpos, Total_cells)
```

## Explore data - plots and summary analysis

### Total cells in crypts

```{r}
# Histogram of total cells
data_processed %>% 
    ggplot(aes(x = Total_cells)) + 
    geom_histogram()

# Different way of generating the same plot
ggplot(data_processed, aes(x = Total_cells)) +
    geom_histogram()

# Histograms of total cells split by mouse
data_processed %>%
    ggplot(aes(x = Total_cells)) + 
    geom_histogram() + 
    facet_wrap(~ Mouse_ID)

# Histograms of total cells split by day post labelling
data_processed %>%
    ggplot(aes(x = Total_cells)) + 
    geom_histogram() +
    facet_wrap(~ Day_post_label)

# Histograms of total cells split by day and sex
data_processed %>% 
    ggplot(aes(x = Total_cells)) + 
    geom_histogram() + 
    facet_grid(Sex ~ Day_post_label)
```

### Labelled cells in crypts

```{r}
# Histogram of RFP positive cells
data_processed %>%
    ggplot(aes(x = RFPpos)) + 
    geom_histogram(bins = 30)

# Histogram of RFP positive cells split by day post label
data_processed %>%
    ggplot(aes(x = RFPpos)) +
    geom_histogram(bins = 20) + 
    facet_wrap(~ Day_post_label)

# Scatter plot of RFP positive cells vs. total cells
data_processed %>% 
    ggplot(aes(x = RFPpos, y = Total_cells, col = Mouse_ID)) + 
    geom_point() + 
    facet_wrap(~ Day_post_label)
```

### Proportion of labelled cells - `mutate()` to add a new column

```{r}
# Add a new column
data_processed <- data_processed %>% 
    mutate(proportion_rfp = RFPpos/Total_cells)

# Histogram of proportion of RFP positive cells
data_processed %>% 
    ggplot(aes(x = proportion_rfp)) + 
    geom_histogram(bins = 20) + 
    facet_wrap(~ Day_post_label)
```

### Summary statistics using `group_by()` and `summarise()`

```{r}
# Summary of numbers                                                        
summary_by_mouse <- data_processed %>% 
    group_by(Mouse_ID) %>% 
    summarise(num_crypts = n(), 
              mean_rfp = mean(RFPpos), 
              mean_num_cells = mean(Total_cells))

summary_by_day <- data_processed %>% 
    group_by(Day_post_label) %>% 
    summarise(num_crypts = n(),
              num_mice = length(unique(Mouse_ID)))
```

### Selecting specific rows (`filter()`) and columns (`select()`)

```{r}
# Keep only large crypts (> 15 cells)
data_processed %>% 
    filter(Total_cells > 15)

# Keep only female
data_processed %>% 
    filter(Sex == "F")

# Keep only mouse 108422-6 
data_processed %>% 
    filter(Mouse_ID == "108422-6")

# Select a few columns
data_processed %>% 
    dplyr::select(Day_post_label:RFPpos)

# Drop a column
data_processed %>% 
    dplyr::select(-Crypt_number)
```
